---
title: "Pilot analysis – clinical & treatment profiles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pilot analysis - clinical & treatment profiles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this report, I run preliminary descriptive regression analyses aiming at describing probability of *"aggmusc"* (**Agg**ressive **Mu**ltiple **Sc**lerosis) given *demographic and clinical* (age at onset, sex, first recorded EDSS, disease duration at the time of the first visit and number of relapses during the first ten years of disease course), and 10-year *treatment profile* (percentage spent treated via Platform, HET and LE HET treatment types) variables.

```{r setup, message=FALSE}
library(aggmusc)
library(dplyr)           # data wrangling
library(ggplot2)         # visualisation
library(marginaleffects) 
```

---

## Data

Data preprocessing process is detailed in the [Data Import vignette](data_import.html).

```{r data, warning=FALSE, message=FALSE, cache=TRUE}
data_path  <- "/Volumes/Extreme SSD/aggrescler/imed_export_20250328_data.xlsx"
valid_path <- "/Volumes/Extreme SSD/aggrescler/patients_with_signed IC_202505_REMUS.xlsx"

data_raw <- prepare_data(data_path, valid_path)

data_classified <- with(data_raw, determine_aggressive_phenotype(
  id, relapses, edss, eye_check = FALSE
))

df <- with(data_raw, preprocess_predictors(
  d0    = data_classified$data,
  edss  = edss,
  treat = treatment,
  rel   = relapses,
  mri   = mri,
  csf   = csf,
  chol  = cholesterol
))

df_model <-
  df |>
  mutate(across(
      c(platform_y10, het_y10, le_het_y10),
      .fns = \(x) if_else(is.na(x), 0, x)
  )) |>
  select(
    id,
    aggmusc         = aggressive,
    sex,
    age_onset,
    dur_first_visit = first_visit_dur,
    edss_first      = first_edss,
    relapse_count,
    platform        = platform_y10,
    het             = het_y10,
    le_het          = le_het_y10
  ) |>
  tidyr::drop_na()
```

## Modelling

```{r modelling}
# ── Model ─────────────────────────────────────────────────────────────────────
fit <- glm(
  aggmusc ~
    age_onset +
    edss_first * dur_first_visit +
    relapse_count +
    splines::ns(platform, 3) +
    het +
    le_het,
  data = df_model,
  family = binomial()
)

# ── Prediction function ───────────────────────────────────────────────────────
make_curve <- function(var, values, df, fit) {
  base <-
    df |>
    summarise(
      age_onset       = median(age_onset),
      dur_first_visit = median(dur_first_visit),
      edss_first      = median(edss_first),
      relapse_count   = median(relapse_count),
      platform        = median(platform),
      het             = median(het),
      le_het          = median(le_het)
    )
    
    grid <- base[rep(1, length(values)), ]
    grid[[var]] <- values
    
    preds <- predictions(
      fit,
      newdata = grid,
      type = "response"
    )
    
    tibble(
      var       = var,
      x         = values,
      pred      = preds$estimate,
      conf_low  = preds$conf.low,
      conf_high = preds$conf.high
    )
}

# ── Build curves ──────────────────────────────────────────────────────────────
curves <- bind_rows(
  make_curve("age_onset",       seq(5, 67, by = 1),   df_model, fit),
  make_curve("dur_first_visit", seq(0, 10, by = 0.5), df_model, fit),
  make_curve("edss_first",      seq(0, 7, by = 0.2),  df_model, fit),
  make_curve("relapse_count",   seq(0, 22, by = 1),   df_model, fit),
  make_curve("platform",        seq(0, 1, by = 0.05), df_model, fit),
  make_curve("het",             seq(0, 1, by = 0.05), df_model, fit),
  make_curve("le_het",          seq(0, 1, by = 0.05), df_model, fit)
)

# ── Long data ─────────────────────────────────────────────────────────────────
df_long <-
  df_model |>
  tidyr::pivot_longer(
    cols = c(age_onset, dur_first_visit, edss_first,
             relapse_count, platform, het, le_het),
    names_to = "var",
    values_to = "value"
  )

# ── Density scale ─────────────────────────────────────────────────────────────
dens_max <-
  df_long |>
  group_by(var) |>
  summarise(
    max_density = max(density(value)$y, na.rm = TRUE),
    .groups = "drop"
  )

# ── Scale curves ──────────────────────────────────────────────────────────────
curves <- 
  curves |>
  left_join(dens_max, by = "var") |>
  group_by(var) |>
  mutate(
    scale_factor     = max_density / max(conf_high, na.rm = TRUE),
    pred_scaled      = pred      * scale_factor,
    conf_low_scaled  = conf_low  * scale_factor,
    conf_high_scaled = conf_high * scale_factor
  ) |>
  ungroup()

# ── Variable names ────────────────────────────────────────────────────────────
var_labels <- c(
  age_onset       = "Age at onset (years)",
  het             = "Time on HET (within 10 years)",
  dur_first_visit = "Disease duration at first visit (years)",
  le_het          = "Time on LE HET (within 10 years)",
  edss_first      = "First EDSS",
  platform        = "Time on Platform (within 10 years)",
  relapse_count   = "Number of relapses (within 10 years)"
)

df_long$var <- factor(
  df_long$var,
  levels = names(var_labels)
)

curves$var <- factor(
  curves$var,
  levels = levels(df_long$var)
)

# ── Second ordinates ──────────────────────────────────────────────────────────
sec_scales <- lapply(levels(curves$var), function(v) {
  df_var <- subset(curves, var == v)
  sf <- unique(df_var$scale_factor)
  
  scale_y_continuous(
    sec.axis = sec_axis(
      trans = \(x) x / sf,
      name = "Predicted probability"
    )
  )
})

# ── Plot ──────────────────────────────────────────────────────────────────────
df_long |>
  ggplot() +
  aes(x = value, fill = factor(aggmusc)) +
  geom_density(alpha = 0.4) +
    
  geom_ribbon(
    data = curves,
    aes(x = x, ymin = conf_low_scaled, ymax = conf_high_scaled),
    inherit.aes = FALSE,
    fill = "red",
    alpha = 0.15
  ) +
    
  geom_line(
    data = curves,
    aes(x = x, y = pred_scaled),
    inherit.aes = FALSE,
    color = "red",
    linewidth = 1
  ) +
  
  facet_wrap(
    ~ var,
    scales = "free",
    ncol = 2,
    labeller = as_labeller(var_labels)
  ) +
  
  ggh4x::facetted_pos_scales(
    y = sec_scales
  ) +
  
  coord_cartesian(ylim = c(0, NA)) +
  
  scale_fill_manual(
    values = c("gray30", "skyblue"),
    labels = c("Non-aggressive", "Aggressive")
  ) +
  
  labs(
    x = "Value",
    y = "Density",
    fill = "Phenotype"
  ) +
    
  theme_classic(base_size = 14) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.spacing = unit(1.2, "lines"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  )
```
