---
title: "Pilot analysis – clinical & treatment profiles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pilot analysis - clinical & treatment profiles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this report, I run preliminary descriptive regression analyses aiming at describing probability of *"aggmusc"* (**Agg**ressive **Mu**ltiple **Sc**lerosis) given *demographic and clinical* (age at onset, sex, first recorded EDSS, disease duration at the time of the first visit and number of relapses during the first ten years of disease course), and 10-year *treatment profile* (percentage spent treated via Platform, HET and LE HET treatment types) variables.

```{r setup, message=FALSE}
library(aggmusc)
library(dplyr)           # data wrangling
library(ggplot2)         # visualisation
library(marginaleffects) # effects computation
```

---

## Data

Data preprocessing process is detailed in the [Data Import vignette](data_import.html).

```{r data, warning=FALSE, message=FALSE, cache=TRUE}
data_path  <- "/Volumes/Extreme SSD/aggrescler/imed_export_20250328_data.xlsx"
valid_path <- "/Volumes/Extreme SSD/aggrescler/patients_with_signed IC_202505_REMUS.xlsx"

data_raw <- prepare_data(data_path, valid_path)

data_classified <- with(data_raw, determine_aggressive_phenotype(
  id, relapses, edss, eye_check = FALSE
))

df <- with(data_raw, preprocess_predictors(
  d0    = data_classified$data,
  edss  = edss,
  treat = treatment,
  rel   = relapses,
  mri   = mri,
  csf   = csf,
  chol  = cholesterol
))

df_model <-
  df |>
  mutate(across(
      c(platform_y10, het_y10, le_het_y10),
      .fns = \(x) if_else(is.na(x), 0, x)
  )) |>
  select(
    id,
    aggmusc                = aggressive,
    sex,
    age_onset,
    dur_first_visit        = first_visit_dur,
    edss_first             = first_edss,
    relapse_count,
    platform               = platform_y10,
    het                    = het_y10,
    le_het                 = le_het_y10,
    dur_first_mri          = first_mri_dur,
    t2_lesions_volume      = t2_lesions_volume_first,
    t1_blackholes_volume   = t1_blackholes_volume_first,
    brain_atrophy          = normalised_brain_atrophy_bpf_sv_first,
    t1_tbv_sv              = t1_tbv_sv_first,
    #corpus_callosum_volume = corpus_callosum_volume_first,
    NULL
  ) |>
  filter(dur_first_mri < 10) |>
  tidyr::drop_na()
```

## Modelling

```{r modelling}
fit_zero <- glm(
  aggmusc ~ 1,
  data = df_model,
  family = binomial()
)

fit_clin <- glm(
  aggmusc ~
    age_onset +
    edss_first +
    dur_first_visit +
    relapse_count +
    platform +
    het +
    le_het,
  data = df_model,
  family = binomial()
)

fit_mri <- glm(
  aggmusc ~
    dur_first_mri +
    t2_lesions_volume +
    t1_blackholes_volume +
    brain_atrophy +
    t1_tbv_sv,
  data = df_model,
  family = binomial()
)

fit_comb <- glm(
  aggmusc ~
    age_onset +
    edss_first +
    dur_first_visit +
    relapse_count +
    platform +
    het +
    le_het +
    dur_first_mri +
    t2_lesions_volume +
    t1_blackholes_volume +
    brain_atrophy +
    t1_tbv_sv,
  data = df_model,
  family = binomial()
)
```

```{r helpers, echo=FALSE}
# ── Prediction function ───────────────────────────────────────────────────────
make_curve <- function(var, values, df, fit) {
  base <-
    df |>
    summarise(
      age_onset            = median(age_onset),
      dur_first_visit      = 2,
      edss_first           = median(edss_first),
      relapse_count        = median(relapse_count),
      platform             = median(platform),
      het                  = median(het),
      le_het               = median(le_het),
      dur_first_mri        = 2,
      t2_lesions_volume    = median(t2_lesions_volume),
      t1_blackholes_volume = median(t1_blackholes_volume),
      brain_atrophy        = median(brain_atrophy),
      t1_tbv_sv            = median(t1_tbv_sv)
    )
    
    grid <- base[rep(1, length(values)), ]
    grid[[var]] <- values
    
    preds <- predictions(
      fit,
      newdata = grid,
      type = "response"
    )
    
    tibble(
      var       = var,
      x         = values,
      pred      = preds$estimate,
      conf_low  = preds$conf.low,
      conf_high = preds$conf.high
    )
}

# ── Jitter binary data ────────────────────────────────────────────────────────
jitt <- function(...) {
  geom_point(
    aes_string(...),
    position = position_jitter(height = 0.05, width = 0.1),
    size = 2,
    shape = 21,
    stroke = 0.4
  )
}

# ── Plot style ────────────────────────────────────────────────────────────────
theme_style <- function() {
  theme_classic(base_size = 12) +
  theme(
    legend.position  = "bottom",
    strip.text       = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.spacing    = unit(1.2, "lines"),
    axis.title       = element_text(face = "bold"),
    legend.title     = element_text(face = "bold")
  )
}
```

```{r fit1-curves, echo=FALSE, warning=FALSE}
# ── Build curves ──────────────────────────────────────────────────────────────
curves <- bind_rows(
  make_curve("age_onset",            seq(5, 67, by = 1),       df_model, fit_comb),
  make_curve("dur_first_visit",      seq(0, 10, by = 0.5),     df_model, fit_comb),
  make_curve("edss_first",           seq(0, 7, by = 0.2),      df_model, fit_comb),
  make_curve("relapse_count",        seq(0, 22, by = 1),       df_model, fit_comb),
  make_curve("platform",             seq(0, 1, by = 0.05),     df_model, fit_comb),
  make_curve("het",                  seq(0, 1, by = 0.05),     df_model, fit_comb),
  make_curve("le_het",               seq(0, 1, by = 0.05),     df_model, fit_comb),
  make_curve("dur_first_mri",        seq(0, 10, by = 0.5),     df_model, fit_comb),
  make_curve("t2_lesions_volume",    seq(0, 30, by = 0.5),     df_model, fit_comb),
  make_curve("t1_blackholes_volume", seq(0, 10, by = 0.5),     df_model, fit_comb),
  make_curve("brain_atrophy",        seq(70, 91, by = 1),      df_model, fit_comb),
  make_curve("t1_tbv_sv",            seq(800, 1700, by = 500), df_model, fit_comb)
)

# ── Long data ─────────────────────────────────────────────────────────────────
df_long <-
  df_model |>
  tidyr::pivot_longer(
    cols = tidyselect::all_of(unique(curves$var)),
    names_to = "var",
    values_to = "value"
  )

# ── Density scale ─────────────────────────────────────────────────────────────
dens_max <-
  df_long |>
  group_by(var) |>
  summarise(
    max_density = max(density(value)$y, na.rm = TRUE),
    .groups = "drop"
  )

# ── Scale curves ──────────────────────────────────────────────────────────────
curves <- 
  curves |>
  left_join(dens_max, by = "var") |>
  group_by(var) |>
  mutate(
    scale_factor     = max_density / max(conf_high, na.rm = TRUE),
    pred_scaled      = pred      * scale_factor,
    conf_low_scaled  = conf_low  * scale_factor,
    conf_high_scaled = conf_high * scale_factor
  ) |>
  ungroup()

# ── Variable names ────────────────────────────────────────────────────────────
var_labels <- c(
  age_onset            = "Age at onset (years)",
  het                  = "Time on HET (within 10 years)",
  dur_first_visit      = "Disease duration at first visit (years)",
  le_het               = "Time on LE HET (within 10 years)",
  edss_first           = "First EDSS",
  platform             = "Time on Platform (within 10 years)",
  relapse_count        = "Number of relapses (within 10 years)",
  dur_first_mri        = "Disease duration at first MRI (years)",
  t2_lesions_volume    = "Lesion volumes (T2)",
  t1_blackholes_volume = "Blackholes volume (T1)",
  brain_atrophy        = "Normalised brain atrophy",
  t1_tbv_sv            = "Total brain volume (T1)"
)

df_long$var <- factor(
  df_long$var,
  levels = names(var_labels)
)

curves$var <- factor(
  curves$var,
  levels = levels(df_long$var)
)

# ── Second ordinates ──────────────────────────────────────────────────────────
sec_scales_clin <- lapply(levels(curves$var), function(v) {
  df_var <- subset(curves, var == v)
  sf <- unique(df_var$scale_factor)
  
  scale_y_continuous(
    sec.axis = sec_axis(
      trans = \(x) x / sf,
      name = "Predicted probability"
    )
  )
})
```

```{r fit1-deer-plot, echo=FALSE, warning=FALSE, fig.height=10}
df_long |>
  ggplot() +
  aes(x = value, fill = factor(aggmusc)) +
  geom_density(alpha = 0.4) +
  geom_ribbon(
    data = curves,
    aes(x = x, ymin = conf_low_scaled, ymax = conf_high_scaled),
    inherit.aes = FALSE,
    fill = "red",
    alpha = 0.15
  ) +
  geom_line(
    data = curves,
    aes(x = x, y = pred_scaled),
    inherit.aes = FALSE,
    color = "red",
    linewidth = 1
  ) +
  facet_wrap(
    ~ var,
    scales = "free",
    ncol = 2,
    labeller = as_labeller(var_labels)
  ) +
  ggh4x::facetted_pos_scales(
    y = sec_scales_clin
  ) +
  coord_cartesian(ylim = c(0, NA)) +
  scale_fill_manual(
    values = c("gray30", "skyblue"),
    labels = c("Non-aggressive", "Aggressive")
  ) +
  labs(
    x = "Value",
    y = "Density",
    fill = "Phenotype"
  ) +
  theme_style()
```

```{r fit1-shark-plot, echo=FALSE, warning=FALSE, fig.height=10}
df_long |>
  ggplot() +
  aes(x = value, y = aggmusc, colour = aggmusc) +
  jitt() +  
  geom_ribbon(
    data = curves,
    aes(x = x, ymin = conf_low, ymax = conf_high),
    inherit.aes = FALSE,
    fill = "red3",
    alpha = 0.15
  ) +
  geom_line(
    data = curves,
    aes(x = x, y = pred),
    inherit.aes = FALSE,
    color = "red3",
    linewidth = 1
  ) +
  facet_wrap(
    ~ var,
    scales = "free_x",
    ncol = 3,
    labeller = as_labeller(var_labels)
  ) +
  scale_y_continuous(breaks = c(0, 0.5, 1)) +
  labs(
    x = "Value",
    y = "Probability of Aggressive Phenotype"
  ) +
  theme_style()
```
