---
title: "Pilot analysis – clinical & treatment profiles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pilot analysis - clinical & treatment profiles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this report, I run preliminary regression analyses aiming at describing
probability of *"aggmusc"* (**Agg**ressive **Mu**ltiple **Sc**lerosis) given
*demographic and clinical* (age at onset, relapses during the first two years
of disease course), 2-year *treatment profile* (percentage spent treated via
Platform, HET and LE HET treatment types), and *MRI* variables.

```{r setup, message=FALSE}
library(aggmusc)
library(tidyverse)       # data wrangling
library(gt)              # tables formatting
library(marginaleffects) # effects computation
library(rstanarm)        # Bayesian modelling
library(projpred)        # variable selection

n_cores <- parallel::detectCores(logical = FALSE)
```

---

## Data and outcome

We analysed a pilot cohort of patients with relapsing multiple sclerosis,
focusing on early clinical and MRI-derived predictors of an aggressive disease
course (hereafter _AggMS_). The binary outcome `aggmusc` indicates whether a
patient meets the criteria for aggressive MS (see
[data import vignette](data_import.html)).

```{r data, warning=FALSE, message=FALSE, cache=TRUE}
data_path  <- "/Volumes/Extreme SSD/aggrescler/imed_export_20250328_data.xlsx"
valid_path <- "/Volumes/Extreme SSD/aggrescler/patients_with_signed IC_202505_REMUS.xlsx"

data_raw <- prepare_data(data_path, valid_path)

data_classified <- with(data_raw, determine_aggressive_phenotype(
  id, relapses, edss, eye_check = FALSE
))

df <- with(data_raw, preprocess_predictors(
  d0    = data_classified$data,
  edss  = edss,
  treat = treatment,
  rel   = relapses,
  mri   = mri,
  csf   = csf,
  chol  = cholesterol
))

df_model <-
  df |>
  mutate(across(
    c(platform_y2, het_y2, le_het_y2),
    .fns = \(x) if_else(is.na(x), 0, x)
  )) |>
  select(
    id,
    aggmusc                = aggressive,
    sex,
    age_onset,
    dur_first_visit        = first_visit_dur,
    edss_first             = first_edss,
    relapse_count          = relapse_count_y2,
    platform               = platform_y2,
    het                    = het_y2,
    le_het                 = le_het_y2,
    dur_first_mri          = first_mri_dur,
    t2_lesions_volume      = t2_lesions_volume_first,
    t1_blackholes_volume   = t1_blackholes_volume_first,
    brain_atrophy          = normalised_brain_atrophy_bpf_sv_first,
    t1_tbv_sv              = t1_tbv_sv_first,
    #corpus_callosum_volume = corpus_callosum_volume_first,
    NULL
  ) |>
  filter(dur_first_mri < 10) |>
  drop_na()

# ── Patients counts ───────────────────────────────────────────────────────────
table(df_model$aggmusc)
```

___

## Frequentist models and incremental predictive value

Predictors were group into three conceptually distinct sets resulting in the
following models:

- **Baseline (M0)**:  
  intercept-only model.
- **Clinical profile (M1)**:  
  age at onset, early relapse activity, and early treatment exposure (time on
  HET / LE-HET / Platform therapy within 2 years).
- **MRI profile (M2)**:  
  disease duration at MRI, total brain volume (T1), normalised brain atrophy,
  lesion volumes (T2), and black hole volumes (T1).
- **Compound model (M3)**:  
  combination of clinical and MRI predictors.

We first estimated a series of logistic regression models and compared them
using likelihood-ratio tests (Type II ANOVA) to assess the incremental
predictive value of clinical and MRI information.

```{r modelling}
# ── Baseline: no predictors ───────────────────────────────────────────────────
fit_zero <- glm(
  aggmusc ~ 1,
  data = df_model,
  family = binomial()
)

# ── Clinical & treatment profile ──────────────────────────────────────────────
fit_clin <- glm(
  aggmusc ~
    age_onset +
    relapse_count +
    platform +
    het +
    le_het,
  data = df_model,
  family = binomial()
)

# ── Magnetic resonance imaging ────────────────────────────────────────────────
fit_mri <- glm(
  aggmusc ~
    dur_first_mri +
    t2_lesions_volume +
    t1_blackholes_volume +
    brain_atrophy +
    t1_tbv_sv,
  data = df_model,
  family = binomial()
)

# ── Combined model ────────────────────────────────────────────────────────────
fit_comb <- glm(
  aggmusc ~
    age_onset +
    relapse_count +
    platform +
    het +
    le_het +
    dur_first_mri +
    t2_lesions_volume +
    t1_blackholes_volume +
    brain_atrophy +
    t1_tbv_sv,
  data = df_model,
  family = binomial()
)
```

___

### Models comparisons

```{r comparisons, echo=FALSE}
comps <- list(
  clin_over_0   = anova(fit_zero, fit_clin),
  mri_over_0    = anova(fit_zero, fit_mri ),
  clin_over_mri = anova(fit_mri,  fit_comb),
  mri_over_clin = anova(fit_clin, fit_comb)
)

model_map <- list(
  "clin_over_0"   = c("M0 (baseline)", "M1 (clinical)", "Predictive value of clinical profile"),
  "mri_over_0"    = c("M0 (baseline)", "M2 (MRI)"     , "Predictive value of MRI profile"),
  "clin_over_mri" = c("M2 (MRI)",      "M3 (compound)", "Predictive value of clinical profile above MRI profile"),
  "mri_over_clin" = c("M1 (clinical)", "M3 (compound)", "Predictive value of MRI profile above clinical profile")
)

extract_anova_data <- function(list) {
  purrr::map_dfr(names(comps), function(i) {
    comp  <- list[[i]]
    stats <- comp[2, ]
    tibble::tibble(
      estimand = model_map[[i]][3],
      model_0  = model_map[[i]][1],
      model_1  = model_map[[i]][2],
      Chisq    = stats[["Deviance"]],
      df       = stats[["Df"]],
      p        = stats[["Pr(>Chi)"]]
    )
  })
}

render_comparison_table <- function(data) {
  data |>
    gt() |>
    cols_label(
      estimand = "Tested effect",
      model_0  = "Model 0",
      model_1  = "Model 1",
      Chisq    = html("&chi;<sup>2</sup>"),
      df       = "df",
      p        = "p"
    ) |>
    fmt_number(
      columns = Chisq,
      decimals = 2
    ) |>
    fmt_number(
      columns = p,
      decimals = 3
    ) |>
    sub_small_vals(
      columns = p,
      threshold = 0.001
    ) |>
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_row_groups()
    )
}

extract_anova_data(comps) |>
  render_comparison_table() |>
  cols_align(align = "center", columns = -1)
```

**Interpretation.**  
Both clinical and MRI predictors substantially improve prediction over baseline.
Importantly, each domain adds unique information beyond the other, as evidenced
by the significant improvements when adding clinical variables to MRI and vice
versa.
___

### Marginal effects

To better understand model behaviour, we inspected marginal effects from the
compound model (M3). Prediction lines for each variable were statistically
fixed at sample median values for all remaining variables (with the exception
of the disease duration at MRI which was statistically fixed at two years).

Results patterns:

- **Age at onset**:  
  Higher age is associated with increased probability of _AggMS_.
- **Disease duration at first MRI**:  
  Weak positive association, suggesting that longer untreated/undocumented
  disease may reflect more aggressive trajectories.
- **Treatment exposure** (within 2 years):  
  Time on HET and LE-HET shows strong mass near zero, limiting interpretability.  
  Time on Platform therapy shows a slight negative association with _AggMS_
  probability.
- **MRI markers**:
  Normalised brain atrophy shows a clear effect: higher atrophy score is
  associated with lower predicted _AggMS_ probability in this dataset.
- **Total brain volume**  
  exhibits a weak negative trend.
- **Lesion volumes (T2)** and **black hole volumes (T1)**  
  show small positive associations.

A key practical observation is the large imbalance in the outcome: most
observations cluster near zero predicted probability, with a small subset near
one. This leads to flat marginal trends for several predictors and wide
uncertainty bands.

```{r helpers, echo=FALSE}
# ── Prediction function ───────────────────────────────────────────────────────
make_curve <- function(var, values, df, fit) {
  base <-
    df |>
    summarise(
      age_onset            = median(age_onset),
      dur_first_visit      = 2,
      edss_first           = median(edss_first),
      relapse_count        = median(relapse_count),
      platform             = median(platform),
      het                  = median(het),
      le_het               = median(le_het),
      dur_first_mri        = 2,
      t2_lesions_volume    = median(t2_lesions_volume),
      t1_blackholes_volume = median(t1_blackholes_volume),
      brain_atrophy        = median(brain_atrophy),
      t1_tbv_sv            = median(t1_tbv_sv)
    )
    
    grid <- base[rep(1, length(values)), ]
    grid[[var]] <- values
    
    preds <- predictions(
      fit,
      newdata = grid,
      type = "response"
    )
    
    tibble(
      var       = var,
      x         = values,
      pred      = preds$estimate,
      conf_low  = preds$conf.low,
      conf_high = preds$conf.high
    )
}

# ── Jitter binary data ────────────────────────────────────────────────────────
jitt <- function(...) {
  geom_point(
    aes_string(...),
    position = position_jitter(height = 0.05, width = 0.1),
    size = 2,
    shape = 21,
    stroke = 0.4
  )
}

# ── Plot style ────────────────────────────────────────────────────────────────
theme_style <- function(leg_pos = "bottom") {
  theme_classic(base_size = 12) +
  theme(
    legend.position  = leg_pos,
    strip.text       = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.spacing    = unit(1.2, "lines"),
    axis.title       = element_text(face = "bold"),
    legend.title     = element_text(face = "bold")
  )
}
```

```{r fit1-curves, echo=FALSE, warning=FALSE}
# ── Build curves ──────────────────────────────────────────────────────────────
curves <- bind_rows(
  make_curve("age_onset",            seq(5, 67, by = 1),       df_model, fit_comb),
  make_curve("relapse_count",        seq(0, 10, by = 0.5),     df_model, fit_comb),
  make_curve("platform",             seq(0, 1, by = 0.05),     df_model, fit_comb),
  make_curve("het",                  seq(0, 1, by = 0.05),     df_model, fit_comb),
  make_curve("le_het",               seq(0, 1, by = 0.05),     df_model, fit_comb),
  make_curve("dur_first_mri",        seq(0, 10, by = 0.5),     df_model, fit_comb),
  make_curve("t2_lesions_volume",    seq(0, 60, by = 1),       df_model, fit_comb),
  make_curve("t1_blackholes_volume", seq(0, 18, by = 1),       df_model, fit_comb),
  make_curve("brain_atrophy",        seq(70, 91, by = 1),      df_model, fit_comb),
  make_curve("t1_tbv_sv",            seq(800, 1700, by = 100), df_model, fit_comb)
)

# ── Long data ─────────────────────────────────────────────────────────────────
df_long <-
  df_model |>
  tidyr::pivot_longer(
    cols = tidyselect::all_of(unique(curves$var)),
    names_to = "var",
    values_to = "value"
  )

# ── Density scale ─────────────────────────────────────────────────────────────
dens_max <-
  df_long |>
  group_by(var) |>
  summarise(
    max_density = max(density(value)$y, na.rm = TRUE),
    .groups = "drop"
  )

# ── Scale curves ──────────────────────────────────────────────────────────────
curves <- 
  curves |>
  left_join(dens_max, by = "var") |>
  group_by(var) |>
  mutate(
    scale_factor     = max_density / max(conf_high, na.rm = TRUE),
    pred_scaled      = pred      * scale_factor,
    conf_low_scaled  = conf_low  * scale_factor,
    conf_high_scaled = conf_high * scale_factor
  ) |>
  ungroup()

# ── Variable names ────────────────────────────────────────────────────────────
var_labels <- c(
  age_onset            = "Age at onset (years)",
  dur_first_mri        = "Disease duration at MRI (years)",
  het                  = "Time on HET (within 2 years)",
  t1_tbv_sv            = "Total brain volume (T1)",
  le_het               = "Time on LE HET (within 2 years)",
  brain_atrophy        = "Normalised brain atrophy",
  platform             = "Time on Platform (within 2 years)",
  t2_lesions_volume    = "Lesion volumes (T2)",
  relapse_count        = "Number of relapses (within 2 years)",
  t1_blackholes_volume = "Blackholes volume (T1)"
)

df_long$var <- factor(
  df_long$var,
  levels = names(var_labels)
)

curves$var <- factor(
  curves$var,
  levels = levels(df_long$var)
)

# ── Second ordinates ──────────────────────────────────────────────────────────
sec_scales_clin <- lapply(levels(curves$var), function(v) {
  df_var <- subset(curves, var == v)
  sf <- unique(df_var$scale_factor)
  
  scale_y_continuous(
    sec.axis = sec_axis(
      trans = \(x) x / sf,
      name = "Predicted probability"
    )
  )
})
```

### Relative effects

The following figure plots data distributions as densities overlaid by model
predictions (red logistic curve with 95% CIs).  
The aim of this plot is to provide visualisation of **relative differences**
between patients with and without _AggMS_, consequently densities are scaled to
maximise contrast between distributions and predicted probabilities plotted on
independent scales (the right ordinate).

```{r shark-plot, echo=FALSE, warning=FALSE, fig.height=10}
df_long |>
  ggplot() +
  aes(x = value, fill = factor(aggmusc)) +
  geom_density(alpha = 0.4) +
  geom_ribbon(
    data = curves,
    aes(x = x, ymin = conf_low_scaled, ymax = conf_high_scaled),
    inherit.aes = FALSE,
    fill = "red",
    alpha = 0.15
  ) +
  geom_line(
    data = curves,
    aes(x = x, y = pred_scaled),
    inherit.aes = FALSE,
    color = "red",
    linewidth = 1
  ) +
  facet_wrap(
    ~ var,
    scales = "free",
    ncol = 2,
    labeller = as_labeller(var_labels)
  ) +
  ggh4x::facetted_pos_scales(
    y = sec_scales_clin
  ) +
  coord_cartesian(ylim = c(0, NA)) +
  scale_fill_manual(
    values = c("gray30", "skyblue"),
    labels = c("Non-aggressive", "Aggressive")
  ) +
  labs(
    x = "Value",
    y = "Density",
    fill = "Phenotype"
  ) +
  theme_style("top")
```

### Absolute effects

The following figure plots data distributions as scatter dots overlaid by model
predictions (red logistic curve with 95% CIs).  
The aim of this plot is to provide visualisation of **absolute differences**
between patients with and without _AggMS_, consequently all ordinates have a
common scale spanning 0-1.

```{r deer-plot, echo=FALSE, warning=FALSE, fig.height=12}
df_long |>
  ggplot() +
  aes(x = value, y = aggmusc, colour = aggmusc) +
  jitt() +  
  geom_ribbon(
    data = curves,
    aes(x = x, ymin = conf_low, ymax = conf_high),
    inherit.aes = FALSE,
    fill = "red3",
    alpha = 0.15
  ) +
  geom_line(
    data = curves,
    aes(x = x, y = pred),
    inherit.aes = FALSE,
    color = "red3",
    linewidth = 1
  ) +
  facet_wrap(
    ~ var,
    scales = "free_x",
    ncol = 2,
    labeller = as_labeller(var_labels)
  ) +
  scale_y_continuous(breaks = c(0, 0.5, 1)) +
  coord_cartesian(ylim = c(-0.05, 1.05)) +
  labs(
    x = "Value",
    y = "Probability of Aggressive Phenotype",
    colour = "Phenotype\n(1 → Aggr.)"
  ) +
  theme_style("right")
```

## Bayesian GLM and projection predictive selection

Next, we estimated a Bayesian logistic regression model using weakly informative
priors. This full model (corresponding to **M3**) serves as the reference model
for projection predictive variable selection (`projpred`).

```{r scale-data}
preds <- c(
  "age_onset",
  "relapse_count",
  "platform",
  "het",
  "le_het",
  "dur_first_mri",
  "t2_lesions_volume",
  "t1_blackholes_volume",
  "brain_atrophy",
  "t1_tbv_sv"
)

scales <- df_model |>
  select(tidyselect::all_of(preds)) |>
  purrr::imap_dfr(function(x, name) {
    tibble(variable = name, M = mean(x), SD = sd(x))
  })

df_model <- df_model |>
  mutate(across(
    tidyselect::all_of(preds),
    .fns = \(x) scale(x)[, 1],
    .names = "{.col}_z"
  ))
```

```{r bayes-mod, message=FALSE, cache=TRUE}
ref_mod <- stan_glm(
  aggmusc ~
    age_onset_z +
    relapse_count_z +
    platform_z +
    het_z +
    le_het_z +
    dur_first_mri_z +
    t2_lesions_volume_z +
    t1_blackholes_volume_z +
    brain_atrophy_z +
    t1_tbv_sv_z,
  data = df_model,
  family = binomial(link = "logit"),
  prior = student_t(df = 7, location = 0, scale = 2.5),
  prior_intercept = student_t(df = 7, location = 0, scale = 2.5),
  cores = n_cores,
  seed = 87542
)
```

### Projection predictive feature selection

Projection predictive selection identified a parsimonious subset of predictors
that retains most of the predictive performance of the full model.

Consistently selected variables included:

- normalised brain atrophy
- age at onset

To the final model we also included disease duration at MRI assessment to allow
for predictions to be compute at specific assessment time values.

```{r preliminary-varsel-fast}
# ── Preliminary fast model ────────────────────────────────────────────────────
refm_obj <- get_refmodel(ref_mod)

cvvs_fast <- cv_varsel(
  refm_obj,
  validate_search = FALSE,
  method = "L1",
  refit_prj = FALSE,
  nterms_max = 20,
  verbose = 0
)

#options(projpred.plot_vsel_size_position = "primary_x_bottom")
#options(projpred.plot_vsel_text_angle = 0)
#plot(cvvs_fast, stats = "mlpd", ranking_nterms_max = NA)

# ── Preliminary refit ─────────────────────────────────────────────────────────
cvvs_fast_refit <- cv_varsel(
  cvvs_fast,
  nclusters_pred = 20,
  verbose = 0
)

#plot(cvvs_fast_refit, stats = "mlpd", ranking_nterms_max = NA)

# ── Final varsel model ────────────────────────────────────────────────────────
cv_fits <- run_cvfun(refm_obj)

use_parallel <- TRUE
if (use_parallel) {
  doParallel::registerDoParallel(n_cores)
}

cvvs <- cv_varsel(
  refm_obj,
  cv_method = "kfold",
  cvfits = cv_fits,
  nclusters_pred = 20,
  parallel = use_parallel,
  verbose = 0
)

if (use_parallel) {
  doParallel::stopImplicitCluster()
  foreach::registerDoSEQ()
}

suggest_size(cvvs, stat = "mlpd")
size_decided <- 2

( rk <- ranking(cvvs) )

( pr_rk <- cv_proportions(rk) )
```

```{r varsel-plot, echo=FALSE}
options(projpred.plot_vsel_show_cv_proportions = TRUE)
plot(cvvs, stats = "mlpd", deltas = TRUE)
```

```{r projected-model, cache=TRUE}
predictors_final <- c(
  head(rk[["fulldata"]], size_decided),
  "dur_first_mri_z" # add MRI date for posterior stratification
)

prj <- project(
  refm_obj,
  predictor_terms = predictors_final,
  verbose = 0
)
```

---

### Posterior predictions

To make the model behaviour more concrete, we computed posterior predictions
from the projection-selected model across a grid of clinically plausible values
of selected variables. Specifically:

- `age_onset`: varied across its observed range
- `brain_atrophy`: varied across its observed range
- `dur_first_mri`: fixed at 2 years

```{r post-pred, echo=FALSE}
# Construct prediction grid
pgrid <- expand_grid(
  age_onset = seq(20, 60, length.out = 25),
  brain_atrophy = seq(70, 90, length.out = 25),
  dur_first_mri = 2
)

# Apply scaling (reuse your scaling table `scls`)
X <- c("age_onset", "brain_atrophy", "dur_first_mri")
pgrid <- pgrid |>
  mutate(across(
    .cols = all_of(X),
    .fns = function(x) {
      scl <- scales |>
        filter(variable == cur_column())
      (x - scl$M) / scl$SD
    },
    .names = "{.col}_scaled"
  ))

# Extract posterior as rvars
b <- posterior::as_draws_rvars(prj)

# Linear predictor
pgrid$linpred <- b$`(Intercept`
for (v in X) {
  pgrid$linpred <- pgrid$linpred +
    b[[paste0(v, "_z")]] * pgrid[[paste0(v, "_scaled")]]
}

# Posterior probabilities
pgrid$prob <- 1 / (1 + exp(-pgrid$linpred))

# Summaries (correct: element-wise over draws)
pgrid$prob_mean <- mean(pgrid$prob)
pgrid$prob_lo   <- posterior::quantile2(pgrid$prob, 0.025)
pgrid$prob_hi   <- posterior::quantile2(pgrid$prob, 0.975)

# Precision (inverse variance)
pgrid$prob_prec    <- 1 / posterior::var(pgrid$prob)
pgrid$alpha_scaled <- scales::rescale(pgrid$prob_prec, to = c(0.5, 1))
#pgrid$alpha_scaled <- scales::rescale(pgrid$prob_hi - pgrid$prob_lo, to = c(1, 0.5))

# Plot
pgrid |>
  ggplot() +
  aes(
    x = brain_atrophy,
    y = age_onset,
    fill = prob_mean,
    alpha = alpha_scaled
  ) +
  geom_tile() +
  scale_fill_viridis_c(
    name = "Pr(AggMS)",
    option = "inferno"
  ) +
  scale_alpha(
    range = c(0.5, 1),
    guide = "none"
  ) +
  labs(
    y = "Age at onset",
    x = "Normalised brain atrophy (at 2 years of disease duration)",
    title = "Posterior predicted probability of Aggressive MS phenotype"
  ) +
  theme_bw()
```

**Interpretation of posterior predictions**  
The posterior prediction surface shows a smooth, low-probability regime across
most of the predictor space, with modest increases in predicted _AggMS_
probability at higher age at onset and lower brain atrophy values
(consistent with the marginal effects).

**Interpretation of posterior predictions**  
The posterior prediction surface shows a strong and approximately monotonic
gradient across the predictor space.

- **Age at onset** shows a clear positive association:  
  higher age is consistently linked to higher predicted probability of _AggMS_
  across all levels of brain atrophy.
- **Normalised brain atrophy** exhibits a pronounced negative association:  
  lower values are associated with substantially higher predicted risk.

These effects combine additively and smoothly, producing a diagonal gradient in
the risk surface:

- The highest predicted probabilities (~0.7–0.85) occur for patients with high
  age at onset and low brain atrophy index.
- The lowest probabilities (<0.1) are concentrated among younger patients with
  high brain atrophy index.
- Intermediate regions show gradual transitions without sharp thresholds,
  indicating that the model does not identify clear decision boundaries but
  rather a continuous risk spectrum.
- Overall, the model shows relatively low prediction precision (indicating by
  desaturated colours across most sections of the plot).

---

## Summary

This pilot analysis examined early clinical, treatment, and MRI predictors of
the aggressive MS phenotype using both frequentist and Bayesian regression
frameworks.

- Likelihood-ratio comparisons showed that **clinical and MRI variables each
  contribute predictive information**, and that their combination improves model
  fit relative to either domain alone.
- Marginal effects from the combined model indicate that **age at onset and
  MRI-derived measures (particularly brain atrophy index)** are the most
  consistently structured predictors. In contrast, treatment exposure variables
  are weakly identified, largely due to their skewed distributions within the
  first two years.
- Projection predictive selection aligns with this pattern: a **low-dimensional
  subset of predictors**, primarily age at onset and brain atrophy, captures most
  of the predictive signal present in the full model.
- Posterior predictions from the projected model show that these predictors
  define a **smooth risk surface spanning a wide range of predicted
  probabilities**. Within this surface:
  - higher age at onset is associated with higher predicted risk,
  - lower brain atrophy index values correspond to higher predicted risk, and
  - their joint effect dominates the model’s behaviour.

Overall, the analysis suggests that:
- **Predictive structure in this dataset is relatively simple**, with most
  signal concentrated in a small number of variables.
- **MRI and clinical information are complementary**, but not all collected
  variables contribute meaningfully once key predictors are accounted for.
- **Treatment variables**, as currently represented, provide **limited
  additional information** in early disease.

Given the pilot nature of the data, the main next steps are:
- **external validation** on an independent cohort, and
- **refinement of predictors**, particularly through
  - improved representation of treatment exposure, and
  - incorporation of longitudinal or more precisely modelled imaging features.
